Ricci-Ollivier Curvature of two Random Walks on the Rooted Phylogenetic Subtree-Prune-Regraft Graph

### 2015-03-30 curvature.tex
Citing Romon.

\subsection{Computing curvature values}
To compute curvature values, we first used \texttt{dense\_spr\_graph} to compute the rSPR graph for four to seven leaves, as discussed in Section~\ref{sec:computing_treespace}.
We then compute curvatures for given pairs of trees directly, by using linear programming to compute the minimal mass transport $W_1$ using the SAGE \cite{SAGE} front-end to the GLPK~\cite{glpk} solver; code can be found in \cite{gricci}.

@@

Next we note a rough bound on the curvature of two trees with respect to their distance, then obtain a tighter bound on the maximum curvature of two adjacent trees.

@@

We continue to explore the effect of biased random walks by considering the curvature of Metropolis Hastings random walks.
The Metropolis Hastings (MH) random walk proposes a move from a tree $T$ to a neighbor tree $S$ uniformly at random and then accepts the move according to the Hastings ratio, $\min\left(1, \frac{\degree{T}}{\degree{S}}\right)$.
The mass distribution for the MH random walk thus leaves a portion of mass at the origin tree, proportional to the relative degree difference of its higher degree neighbors.
We now bound the difference between the curvature of uniform random walks $\curvature{S,T}$, and that of MH random walks, $\curvature[\MH]{S,T}$.

>>>>

\subsection{Computing curvature values}
To compute curvature values, we first used \texttt{dense\_spr\_graph} to compute the rSPR graph for four to seven leaves, as discussed in Section~\ref{sec:computing_treespace}.
We then compute curvatures for given pairs of trees directly, by using linear programming to compute the minimal mass transport $W_1$ using the SAGE \cite{SAGE} front-end to the GLPK~\cite{glpk} solver; code can be found in \cite{gricci} which grew from the code described in \cite{Loisel2014-hu}.

@@

Next we note a simple and rough bound on the curvature of two trees with respect to their distance, then obtain a tighter bound on the maximum curvature of two adjacent trees.

@@

Next we consider the curvature of the Metropolis-Hastings (MH) random walk sampling the uniform distribution on trees, which we will simply call the MH random walk.
This random walk proposes a move from a tree $T$ to a neighbor tree $S$ uniformly at random and then accepts the move according to the Hastings ratio, $\min\left(1, \frac{\degree{T}}{\degree{S}}\right)$.
The mass distribution for the MH random walk thus leaves a portion of mass at the origin tree, proportional to the relative degree difference of its higher degree neighbors.
We now bound the difference between the curvature of uniform random walks $\curvature{S,T}$, and that of MH random walks, $\curvature[\MH]{S,T}$.
